---
title: "Nanopore IGROV-1 bulk RNA-Seq and scRNA-Seq correlation benchmarks"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
library(SingleCellExperiment)
```

Global parameters:
```{r}
bulk_dir <- file.path("..", "nanopore_bulk_analysis", "report_data")
sc_dir <- file.path("..", "nanopore_sc_analysis", "report_data")
souporcell_file <- file.path("..", "illumina_sc_analysis", "data",
                             "souporcell_clusters.tsv")
bulk_sample_ids <- c("LIB5432315_SAM24385458", "LIB5432316_SAM24385459",
                     "LIB5427896_SAM24376275", "LIB5427897_SAM24376276")
k_values <- c(1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 800)
min_tpm <- 1
top_n_transctipts <- 5000
n_values <- c(100, 200, 500, 1000, 2000, 5000, 10000, 15000, 20000)
```

Helper functions:
```{r}
fill_missing_matrix <- function(x, all_rownames) {
    missing_rownames <- setdiff(all_rownames, rownames(x))
    missing_matrix <- matrix(
        0, nrow = length(missing_rownames), ncol = ncol(x)
    )
    rownames(missing_matrix) <- missing_rownames
    full_matrix <- rbind(x, missing_matrix)
    full_matrix <- full_matrix[all_rownames,]
    return(full_matrix)
}
calculate_bulk_pseudobulk_cor <- function(bulk_tpm_list, pseudobulk_tpm,
                                          bulk_name, k_values, method) {
    cor_values <- sapply(k_values, function(k_value) {
        bulk_tpm_values <- bulk_tpm_list[[bulk_name]]
        bulk_tpm_values <- head(
            sort(bulk_tpm_values, decreasing = TRUE),
            n = top_n_transctipts
        )
        cor(bulk_tpm_values,
            pseudobulk_tpm[names(bulk_tpm_values), as.character(k_value)],
            method = method)
    })
    names(cor_values) = k_values
    return(cor_values)
}
cor_top_n <- function(bulk_tpm_list, pseudobulk_tpm,
                      bulk_name, k_value, n_values, method) {
    bulk_tpm_values <- bulk_tpm_list[[bulk_name]]
    bulk_tpm_values <- sort(bulk_tpm_values, decreasing = TRUE)
    pseudobulk_tpm_values <- pseudobulk_tpm[
        names(bulk_tpm_values), as.character(k_value)
    ]
    cor_values <- sapply(n_values, function(n) {
        if (length(bulk_tpm_values) < n) {
            return(NA)
        } else {
            return(cor(bulk_tpm_values[1:n],
                       pseudobulk_tpm_values[1:n],
                       method = method))
        }
    })
    return(cor_values)
}
```

Prepare mitochondrial transcript blacklist:
```{r, message = FALSE, warning = FALSE}
se_transcript_mt <- readRDS(file.path(
    sc_dir, "isosceles_se_igrov_pseudobulk_k_transcript.rds"
))
se_transcript_mt <- se_transcript_mt[
    grepl("^MT:", rowData(se_transcript_mt)$position),
]
mt_transcripts_isosceles <- rownames(se_transcript_mt)
mt_transcripts_ensembl <- rowData(se_transcript_mt)$compatible_tx
mt_transcripts_isosceles <- c(
    mt_transcripts_isosceles,
    c("ISOT-0000-0000-0000-2622:s10197250:e10199150:FA:FL",
      "ISOT-0000-0000-0000-caeb:s10507850:e10509200:FA:FL",
      "ISOT-0000-0000-0000-2809:s96617150:e96618250:FA:FL")
)
mt_transcripts_ensembl <- c(
    mt_transcripts_ensembl,
    c("ENST00000445125", "ENST00000536684", "ENST00000600213")
)
```

Prepare Isosceles bulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
isosceles_bulk_tpm <- sapply(bulk_sample_ids, function(sample_id) {
    se_transcript <- readRDS(file.path(
        bulk_dir, glue("isosceles_{sample_id}_se_transcript.rds")
    ))
    return(assay(se_transcript, "tpm")[, 1])
})
isosceles_bulk_tpm <- isosceles_bulk_tpm[
    !(rownames(isosceles_bulk_tpm) %in% mt_transcripts_isosceles),
]
isosceles_bulk_promethion_tpm <- apply(isosceles_bulk_tpm[, 1:2], 1, mean)
isosceles_bulk_promethion_tpm <- isosceles_bulk_promethion_tpm[
    isosceles_bulk_promethion_tpm >= min_tpm
]
isosceles_bulk_minion_tpm <- apply(isosceles_bulk_tpm[, 3:4], 1, mean)
isosceles_bulk_minion_tpm <- isosceles_bulk_minion_tpm[
    isosceles_bulk_minion_tpm >= min_tpm
]
isosceles_bulk_tpm_list <- list(
    Promethion = isosceles_bulk_promethion_tpm,
    MinION = isosceles_bulk_minion_tpm
)
```

Prepare Isosceles pseudobulk RNA-Seq data (k cells subsets):
```{r, message = FALSE, warning = FALSE}
isosceles_pseudobulk_tpm <- assay(readRDS(file.path(
    sc_dir, "isosceles_se_igrov_pseudobulk_k_transcript.rds"
)), "tpm")
isosceles_pseudobulk_tpm <- isosceles_pseudobulk_tpm[
    !(rownames(isosceles_pseudobulk_tpm) %in% mt_transcripts_isosceles),
]
```

Calculate Isosceles correlations:
```{r, message = FALSE, warning = FALSE}
isosceles_cor_df <- data.frame(
    tool = "Isosceles",
    platform = rep(c("Promethion", "MinION"), each = length(k_values) * 2),
    cor_type = rep(c("pearson", "spearman"), each = length(k_values), 2),
    k_value = rep(k_values, 4),
    cor_value = c(
        calculate_bulk_pseudobulk_cor(
            isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
            "Promethion", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
            "Promethion", k_values, "spearman"
        ),
        calculate_bulk_pseudobulk_cor(
            isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
            "MinION", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
            "MinION", k_values, "spearman"
        )
    )
)
```

Prepare Souporcell cluster data:
```{r, message = FALSE, warning = FALSE}
souporcell_df <- read.delim(souporcell_file)
souporcell_barcode <- gsub("-1$", "", souporcell_df$barcode)
souporcell_cluster <- ifelse(souporcell_df$status == "singlet",
                             souporcell_df$assignment,
                             souporcell_df$status)
souporcell_cluster <- setNames(souporcell_cluster, souporcell_barcode)
```

Prepare IsoQuant bulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
isoquant_bulk_list <- lapply(bulk_sample_ids, function(sample_id) {
    isoquant_df <- read_delim(file.path(
        bulk_dir, glue("isoquant_{sample_id}_transcript_tpm.tsv")
    ))
})
isoquant_bulk_tx_ids <- unique(unlist(sapply(
    isoquant_bulk_list, function(df) {df[, 1, drop = TRUE]}
)))
isoquant_bulk_tpm <- sapply(isoquant_bulk_list, function(df) {
    tpm_values <- df$TPM
    names(tpm_values) <- df[, 1, drop = TRUE]
    tpm_values <- tpm_values[isoquant_bulk_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(isoquant_bulk_tpm) <- isoquant_bulk_tx_ids
colnames(isoquant_bulk_tpm) <- bulk_sample_ids
isoquant_bulk_tpm <- isoquant_bulk_tpm[
    !(rownames(isoquant_bulk_tpm) %in% mt_transcripts_ensembl),
]
isoquant_bulk_promethion_tpm <- apply(isoquant_bulk_tpm[, 1:2], 1, mean)
isoquant_bulk_promethion_tpm <- isoquant_bulk_promethion_tpm[
    isoquant_bulk_promethion_tpm >= min_tpm
]
isoquant_bulk_minion_tpm <- apply(isoquant_bulk_tpm[, 3:4], 1, mean)
isoquant_bulk_minion_tpm <- isoquant_bulk_minion_tpm[
    isoquant_bulk_minion_tpm >= min_tpm
]
isoquant_bulk_tpm_list <- list(
    Promethion = isoquant_bulk_promethion_tpm,
    MinION = isoquant_bulk_minion_tpm
)
```

Prepare IsoQuant scRNA-Seq data:
```{r, message = FALSE, warning = FALSE}
isoquant_sc_df <- read_delim(file.path(
    sc_dir, "isoquant_transcript_grouped_tpm.tsv"
))
isoquant_sc_transcript_ids <- isoquant_sc_df[, 1, drop = TRUE]
isoquant_sc_counts <- as(as.matrix(isoquant_sc_df[, c(-1)]), "dgCMatrix")
rownames(isoquant_sc_counts) <- isoquant_sc_transcript_ids
isoquant_sc_counts <- isoquant_sc_counts[
    rownames(isoquant_sc_counts) %in% rownames(isoquant_bulk_tpm),
]
cell_labels <- souporcell_cluster[colnames(isoquant_sc_counts)]
cell_selector <- cell_labels == "1"
isoquant_sc_counts <- isoquant_sc_counts[, cell_selector]
cell_order <- order(colSums(isoquant_sc_counts), decreasing = TRUE)
isoquant_sc_counts <- isoquant_sc_counts[, cell_order]
```

Prepare IsoQuant pseudobulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
isoquant_pseudobulk_tpm <- lapply(k_values, function(k_value) {
    rowSums(isoquant_sc_counts[, 1:k_value, drop = FALSE])
})
isoquant_pseudobulk_tpm <- do.call(cbind, isoquant_pseudobulk_tpm)
colnames(isoquant_pseudobulk_tpm) <- k_values
isoquant_pseudobulk_tpm <- fill_missing_matrix(isoquant_pseudobulk_tpm,
                                               rownames(isoquant_bulk_tpm))
isoquant_pseudobulk_tpm <- t(
    t(isoquant_pseudobulk_tpm) / colSums(isoquant_pseudobulk_tpm) * 1e6
)
```

Calculate IsoQuant correlations:
```{r, message = FALSE, warning = FALSE}
isoquant_cor_df <- data.frame(
    tool = "IsoQuant",
    platform = rep(c("Promethion", "MinION"), each = length(k_values) * 2),
    cor_type = rep(c("pearson", "spearman"), each = length(k_values), 2),
    k_value = rep(k_values, 4),
    cor_value = c(
        calculate_bulk_pseudobulk_cor(
            isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
            "Promethion", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
            "Promethion", k_values, "spearman"
        ),
        calculate_bulk_pseudobulk_cor(
            isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
            "MinION", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
            "MinION", k_values, "spearman"
        )
    )
)
```

Prepare FLAMES bulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
flames_bulk_list <- lapply(bulk_sample_ids, function(sample_id) {
    flames_df <- read_csv(file.path(
        bulk_dir, glue("flames_{sample_id}_transcript_count.csv.gz")
    ))
    flames_df <- flames_df[!grepl("_", flames_df$transcript_id),]
    return(flames_df)
})
flames_bulk_tx_ids <- unique(unlist(sapply(
    flames_bulk_list, function(df) {df$transcript_id}
)))
flames_bulk_tpm <- sapply(flames_bulk_list, function(df) {
    tpm_values <- df[, 3, drop = TRUE]
    names(tpm_values) <- df$transcript_id
    tpm_values <- tpm_values[flames_bulk_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(flames_bulk_tpm) <- flames_bulk_tx_ids
colnames(flames_bulk_tpm) <- bulk_sample_ids
flames_bulk_tpm <- t(t(flames_bulk_tpm) / colSums(flames_bulk_tpm) * 1e6)
flames_bulk_tpm <- flames_bulk_tpm[
    !(rownames(flames_bulk_tpm) %in% mt_transcripts_ensembl),
]
flames_bulk_promethion_tpm <- apply(flames_bulk_tpm[, 1:2], 1, mean)
flames_bulk_promethion_tpm <- flames_bulk_promethion_tpm[
    flames_bulk_promethion_tpm >= min_tpm
]
flames_bulk_minion_tpm <- apply(flames_bulk_tpm[, 3:4], 1, mean)
flames_bulk_minion_tpm <- flames_bulk_minion_tpm[
    flames_bulk_minion_tpm >= min_tpm
]
flames_bulk_tpm_list <- list(
    Promethion = flames_bulk_promethion_tpm,
    MinION = flames_bulk_minion_tpm
)
```

Prepare FLAMES scRNA-Seq data:
```{r, message = FALSE, warning = FALSE}
flames_sc_df <- read_csv(file.path(
    sc_dir, "flames_transcript_count.csv.gz"
))
flames_sc_transcript_ids <- flames_sc_df$transcript_id
flames_sc_counts <- as(as.matrix(flames_sc_df[, c(-1, -2)]), "dgCMatrix")
rownames(flames_sc_counts) <- flames_sc_transcript_ids
flames_sc_counts <- flames_sc_counts[
    !grepl("_", rownames(flames_sc_counts)),
]
flames_sc_counts <- flames_sc_counts[
    rownames(flames_sc_counts) %in% rownames(flames_bulk_tpm),
]
cell_labels <- souporcell_cluster[colnames(flames_sc_counts)]
cell_selector <- cell_labels == "1"
flames_sc_counts <- flames_sc_counts[, cell_selector]
cell_order <- order(colSums(flames_sc_counts), decreasing = TRUE)
flames_sc_counts <- flames_sc_counts[, cell_order]
```

Prepare FLAMES pseudobulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
flames_pseudobulk_tpm <- lapply(k_values, function(k_value) {
    rowSums(flames_sc_counts[, 1:k_value, drop = FALSE])
})
flames_pseudobulk_tpm <- do.call(cbind, flames_pseudobulk_tpm)
colnames(flames_pseudobulk_tpm) <- k_values
flames_pseudobulk_tpm <- fill_missing_matrix(flames_pseudobulk_tpm,
                                             rownames(flames_bulk_tpm))
flames_pseudobulk_tpm <- t(
    t(flames_pseudobulk_tpm) / colSums(flames_pseudobulk_tpm) * 1e6
)
```

Calculate FLAMES correlations:
```{r, message = FALSE, warning = FALSE}
flames_cor_df <- data.frame(
    tool = "FLAMES",
    platform = rep(c("Promethion", "MinION"), each = length(k_values) * 2),
    cor_type = rep(c("pearson", "spearman"), each = length(k_values), 2),
    k_value = rep(k_values, 4),
    cor_value = c(
        calculate_bulk_pseudobulk_cor(
            flames_bulk_tpm_list, flames_pseudobulk_tpm,
            "Promethion", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            flames_bulk_tpm_list, flames_pseudobulk_tpm,
            "Promethion", k_values, "spearman"
        ),
        calculate_bulk_pseudobulk_cor(
            flames_bulk_tpm_list, flames_pseudobulk_tpm,
            "MinION", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            flames_bulk_tpm_list, flames_pseudobulk_tpm,
            "MinION", k_values, "spearman"
        )
    )
)
```

Prepare Sicelore bulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
sicelore_bulk_list <- lapply(bulk_sample_ids, function(sample_id) {
    sicelore_df <- read_delim(file.path(
        bulk_dir, glue("sicelore_{sample_id}_sicelore_isomatrix.txt")
    ))
    sicelore_df <- sicelore_df[sicelore_df$transcriptId != "undef",]
    return(sicelore_df)
})
sicelore_bulk_tx_ids <- unique(unlist(sapply(
    sicelore_bulk_list, function(df) {df$transcriptId}
)))
sicelore_bulk_tpm <- sapply(sicelore_bulk_list, function(df) {
    tpm_values <- df[, 4, drop = TRUE]
    names(tpm_values) <- df$transcriptId
    tpm_values <- tpm_values[sicelore_bulk_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(sicelore_bulk_tpm) <- sicelore_bulk_tx_ids
colnames(sicelore_bulk_tpm) <- bulk_sample_ids
sicelore_bulk_tpm <- t(
    t(sicelore_bulk_tpm) / colSums(sicelore_bulk_tpm) * 1e6
)
sicelore_bulk_tpm <- sicelore_bulk_tpm[
    !(rownames(sicelore_bulk_tpm) %in% mt_transcripts_ensembl),
]
sicelore_bulk_promethion_tpm <- apply(sicelore_bulk_tpm[, 1:2], 1, mean)
sicelore_bulk_promethion_tpm <- sicelore_bulk_promethion_tpm[
    sicelore_bulk_promethion_tpm >= min_tpm
]
sicelore_bulk_minion_tpm <- apply(sicelore_bulk_tpm[, 3:4], 1, mean)
sicelore_bulk_minion_tpm <- sicelore_bulk_minion_tpm[
    sicelore_bulk_minion_tpm >= min_tpm
]
sicelore_bulk_tpm_list <- list(
    Promethion = sicelore_bulk_promethion_tpm,
    MinION = sicelore_bulk_minion_tpm
)
```

Prepare Sicelore scRNA-Seq data:
```{r, message = FALSE, warning = FALSE}
sicelore_sc_df <- read_delim(file.path(
    sc_dir, "sicelore_isomatrix.txt"
))
sicelore_sc_transcript_ids <- sicelore_sc_df$transcriptId
sicelore_sc_counts <- as(as.matrix(sicelore_sc_df[, c(-1, -2, -3)]),
                         "dgCMatrix")
rownames(sicelore_sc_counts) <- sicelore_sc_transcript_ids
sicelore_sc_counts <- sicelore_sc_counts[
    rownames(sicelore_sc_counts) != "undef",
]
sicelore_sc_counts <- sicelore_sc_counts[
    rownames(sicelore_sc_counts) %in% rownames(sicelore_bulk_tpm),
]
cell_labels <- souporcell_cluster[colnames(sicelore_sc_counts)]
cell_selector <- cell_labels == "1"
sicelore_sc_counts <- sicelore_sc_counts[, cell_selector]
cell_order <- order(colSums(sicelore_sc_counts), decreasing = TRUE)
sicelore_sc_counts <- sicelore_sc_counts[, cell_order]
```

Prepare Sicelore pseudobulk RNA-Seq data:
```{r, message = FALSE, warning = FALSE}
sicelore_pseudobulk_tpm <- lapply(k_values, function(k_value) {
    rowSums(sicelore_sc_counts[, 1:k_value, drop = FALSE])
})
sicelore_pseudobulk_tpm <- do.call(cbind, sicelore_pseudobulk_tpm)
colnames(sicelore_pseudobulk_tpm) <- k_values
sicelore_pseudobulk_tpm <- fill_missing_matrix(sicelore_pseudobulk_tpm,
                                               rownames(sicelore_bulk_tpm))
sicelore_pseudobulk_tpm <- t(
    t(sicelore_pseudobulk_tpm) / colSums(sicelore_pseudobulk_tpm) * 1e6
)
```

Calculate Sicelore correlations:
```{r, message = FALSE, warning = FALSE}
sicelore_cor_df <- data.frame(
    tool = "Sicelore",
    platform = rep(c("Promethion", "MinION"), each = length(k_values) * 2),
    cor_type = rep(c("pearson", "spearman"), each = length(k_values), 2),
    k_value = rep(k_values, 4),
    cor_value = c(
        calculate_bulk_pseudobulk_cor(
            sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
            "Promethion", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
            "Promethion", k_values, "spearman"
        ),
        calculate_bulk_pseudobulk_cor(
            sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
            "MinION", k_values, "pearson"
        ),
        calculate_bulk_pseudobulk_cor(
            sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
            "MinION", k_values, "spearman"
        )
    )
)
```

# Plots

Prepare correlation data:
```{r, message = FALSE, warning = FALSE}
cor_df <- rbind(isosceles_cor_df, isoquant_cor_df, flames_cor_df,
                sicelore_cor_df)
```

Helper function:
```{r}
plot_cor <- function(method) {
    plot_df <- filter(cor_df, cor_type == method)
    ggplot(plot_df, mapping = aes(x = k_value,
                                  y = cor_value,
                                  col = tool,
                                  linetype = platform)) +
        geom_line(size = 1) +
        scale_color_manual(values = c(Isosceles = "black",
                                      IsoQuant = "gold",
                                      FLAMES = "tomato",
                                      Sicelore = "royalblue")) +
        scale_linetype_manual(values = c(Promethion = "solid",
                                         MinION = "dotted")) +
        scale_x_continuous(
            trans = "log2",
            breaks = c(1, 4, 16, 64, 256, 800)
        ) +
        labs(
           x = "Top k cells",
           y = "Correlation value",
           col = "Tool",
           linetype = "Platform"
        ) +
        theme_bw() +
        theme(aspect.ratio = 1,
              legend.title = element_blank())
}
plot_cor_top_n <- function(method, k_value, n_values, ylim) {
    plot_df <- list(
        data.frame(
            tool = "Isosceles",
            platform = rep(c("Promethion", "MinION"),
                           each = length(n_values)),
            n_value = rep(n_values, 2),
            cor_value = c(
                cor_top_n(
                    isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
                    "Promethion", k_value, n_values, method
                ),
                cor_top_n(
                    isosceles_bulk_tpm_list, isosceles_pseudobulk_tpm,
                    "MinION", k_value, n_values, method
                )
            )
        ),
        data.frame(
            tool = "IsoQuant",
            platform = rep(c("Promethion", "MinION"),
                           each = length(n_values)),
            n_value = rep(n_values, 2),
            cor_value = c(
                cor_top_n(
                    isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
                    "Promethion", k_value, n_values, method
                ),
                cor_top_n(
                    isoquant_bulk_tpm_list, isoquant_pseudobulk_tpm,
                    "MinION", k_value, n_values, method
                )
            )
        ),
        data.frame(
            tool = "FLAMES",
            platform = rep(c("Promethion", "MinION"),
                           each = length(n_values)),
            n_value = rep(n_values, 2),
            cor_value = c(
                cor_top_n(
                    flames_bulk_tpm_list, flames_pseudobulk_tpm,
                    "Promethion", k_value, n_values, method
                ),
                cor_top_n(
                    flames_bulk_tpm_list, flames_pseudobulk_tpm,
                    "MinION", k_value, n_values, method
                )
            )
        ),
        data.frame(
            tool = "Sicelore",
            platform = rep(c("Promethion", "MinION"),
                           each = length(n_values)),
            n_value = rep(n_values, 2),
            cor_value = c(
                cor_top_n(
                    sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
                    "Promethion", k_value, n_values, method
                ),
                cor_top_n(
                    sicelore_bulk_tpm_list, sicelore_pseudobulk_tpm,
                    "MinION", k_value, n_values, method
                )
            )
        )
    )
    plot_df <- do.call(rbind, plot_df)
    plot_df <- plot_df[!is.na(plot_df$cor_value),]
    ggplot(plot_df, mapping = aes(x = n_value,
                                  y = cor_value,
                                  col = tool,
                                  linetype = platform)) +
        geom_line(size = 1) +
        scale_color_manual(values = c(Isosceles = "black",
                                      IsoQuant = "gold",
                                      FLAMES = "tomato",
                                      Sicelore = "royalblue")) +
        scale_linetype_manual(values = c(Promethion = "solid",
                                         MinION = "dotted")) +
        scale_x_continuous(
            trans = "log10",
            n.breaks = 7
        ) +
        lims(y = ylim) +
        labs(
           x = "Top N transcripts",
           y = "Correlation value",
           col = "Tool",
           linetype = "Platform"
        ) +
        theme_bw() +
        theme(aspect.ratio = 1,
              legend.title = element_blank())
}
```

Bulk vs pseudobulk Spearmancorrelation plot (top 5000 transcripts):
```{r, message = FALSE, warning = FALSE}
plot_cor("spearman")
```

Bulk vs pseudobulk Spearman correlation plot for top N transcripts (top 64 cells):
```{r, message = FALSE, warning = FALSE}
plot_cor_top_n("spearman", 64, n_values, c(0.1, 0.8))
```

Bulk vs pseudobulk Spearman correlation plot for top N transcripts (top 800 cells):
```{r, message = FALSE, warning = FALSE}
plot_cor_top_n("spearman", 800, n_values, c(0.1, 0.8))
```

# Session Info

```{r}
sessionInfo()
```
