---
title: "Nanopore IGROV-1 bulk RNA-Seq mean relative difference benchmarks"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
library(SingleCellExperiment)
```

Global parameters:
```{r}
result_dir <- file.path("..", "nanopore_bulk_analysis", "report_data")
sample_ids <- c("LIB5432315_SAM24385458", "LIB5432316_SAM24385459",
                "LIB5427896_SAM24376275", "LIB5427897_SAM24376276")
min_tpm <- 1
```

Helper functions:
```{r}
mean_rel_diff <- function(x, y) {
    mean(abs(x - y) / ((x + y) / 2))
}
calculate_mean_rel_diff <- function(tpm_data) {
    promethion_tpm <- tpm_data[, 1:2]
    minion_tpm <- tpm_data[, 3:4]
    mean_tpm <- cbind(
        rowSums(promethion_tpm) / 2,
        rowSums(minion_tpm) / 2
    )
    promethion_tpm <- promethion_tpm[
        apply(promethion_tpm, 1, mean) >= min_tpm,
    ]
    minion_tpm <- minion_tpm[
        apply(minion_tpm, 1, mean) >= min_tpm,
    ]
    mean_tpm <- mean_tpm[
        apply(mean_tpm, 1, mean) >= min_tpm,
    ]
    print(paste0("Filtered transcripts (Promethion vs Promethion): ",
                 nrow(promethion_tpm)))
    print(paste0("Filtered transcripts (MinION vs MinION): ",
                 nrow(minion_tpm)))
    print(paste0("Filtered transcripts (MinION vs Promethion): ",
                 nrow(mean_tpm)))
    mean_rel_diff_vector <- c(
        MM = mean_rel_diff(minion_tpm[, 1], minion_tpm[, 2]),
        PP = mean_rel_diff(promethion_tpm[, 1], promethion_tpm[, 2]),
        MP = mean_rel_diff(mean_tpm[, 1], mean_tpm[, 2])
    )
    return(mean_rel_diff_vector)
}
```

Prepare Isosceles relative difference data:
```{r, message = FALSE, warning = FALSE}
isosceles_tpm <- sapply(sample_ids, function(sample_id) {
    se_transcript <- readRDS(file.path(
        result_dir, glue("isosceles_{sample_id}_se_transcript.rds")
    ))
    return(assay(se_transcript, "tpm")[, 1])
})
isosceles_mean_rel_diff <- calculate_mean_rel_diff(isosceles_tpm)
```

Prepare IsoQuant relative difference data:
```{r, message = FALSE, warning = FALSE}
isoquant_list <- lapply(sample_ids, function(sample_id) {
    isoquant_df <- read_delim(file.path(
        result_dir, glue("isoquant_{sample_id}_transcript_tpm.tsv")
    ))
})
isoquant_tx_ids <- unique(unlist(sapply(
    isoquant_list, function(df) {df[, 1, drop = TRUE]}
)))
isoquant_tpm <- sapply(isoquant_list, function(df) {
    tpm_values <- df$TPM
    names(tpm_values) <- df[, 1, drop = TRUE]
    tpm_values <- tpm_values[isoquant_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(isoquant_tpm) <- isoquant_tx_ids
colnames(isoquant_tpm) <- sample_ids
isoquant_mean_rel_diff <- calculate_mean_rel_diff(isoquant_tpm)
```

Prepare bambu relative difference data:
```{r, message = FALSE, warning = FALSE}
bambu_tpm <- sapply(sample_ids, function(sample_id) {
    bambu_df <- read_delim(file.path(
        result_dir, glue("bambu_{sample_id}_counts_transcript.txt")
    ))
    return(setNames(bambu_df[, 3, drop = TRUE], bambu_df$TXNAME))
})
bambu_tpm <- t(t(bambu_tpm) / colSums(bambu_tpm) * 1e6)
bambu_mean_rel_diff <- calculate_mean_rel_diff(bambu_tpm)
```

Prepare Flair relative difference data:
```{r, message = FALSE, warning = FALSE}
flair_list <- lapply(sample_ids, function(sample_id) {
    flair_df <- read_delim(file.path(
        result_dir, glue("flair_{sample_id}_counts_matrix.tsv")
    ))
    flair_df <- flair_df[grepl("^ENST", flair_df$ids),]
    flair_df <- flair_df[!grepl("-", flair_df$ids),]
})
flair_tx_ids <- unique(unlist(sapply(
    flair_list, function(df) {df$ids}
)))
flair_tpm <- sapply(flair_list, function(df) {
    tpm_values <- df$sample_condition_batch
    names(tpm_values) <- df$ids
    tpm_values <- tpm_values[flair_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(flair_tpm) <- sapply(strsplit(flair_tx_ids, "_"), "[", 1)
colnames(flair_tpm) <- sample_ids
flair_tpm <- t(t(flair_tpm) / colSums(flair_tpm) * 1e6)
flair_mean_rel_diff <- calculate_mean_rel_diff(flair_tpm)
```

Prepare NanoCount relative difference data:
```{r, message = FALSE, warning = FALSE}
nanocount_list <- lapply(sample_ids, function(sample_id) {
    nanocount_df <- read_delim(file.path(
        result_dir, glue("nanocount_{sample_id}.tsv")
    ))
})
nanocount_tx_ids <- unique(unlist(sapply(
    nanocount_list, function(df) {df$transcript_name}
)))
nanocount_tpm <- sapply(nanocount_list, function(df) {
    tpm_values <- df$tpm
    names(tpm_values) <- df$transcript_name
    tpm_values <- tpm_values[nanocount_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(nanocount_tpm) <- nanocount_tx_ids
colnames(nanocount_tpm) <- sample_ids
nanocount_mean_rel_diff <- calculate_mean_rel_diff(nanocount_tpm)
```

Prepare LIQA relative difference data:
```{r, message = FALSE, warning = FALSE}
liqa_list <- lapply(sample_ids, function(sample_id) {
    liqa_df <- read_delim(file.path(
        result_dir, glue("liqa_{sample_id}.tsv")
    ))
})
liqa_tx_ids <- unique(unlist(sapply(
    liqa_list, function(df) {df$IsoformName}
)))
liqa_tpm <- sapply(liqa_list, function(df) {
    tpm_values <- df$ReadPerGene_corrected
    names(tpm_values) <- df$IsoformName
    tpm_values <- tpm_values[liqa_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(liqa_tpm) <- liqa_tx_ids
colnames(liqa_tpm) <- sample_ids
liqa_tpm <- t(t(liqa_tpm) / colSums(liqa_tpm) * 1e6)
liqa_mean_rel_diff <- calculate_mean_rel_diff(liqa_tpm)
```

Prepare ESPRESSO relative difference data:
```{r, message = FALSE, warning = FALSE}
espresso_list <- lapply(sample_ids, function(sample_id) {
    espresso_df <- read_delim(file.path(
        result_dir, glue("espresso_{sample_id}_abundance.esp")
    ))
    espresso_df <- espresso_df[grepl("^ENST", espresso_df$transcript_ID),]
})
espresso_tx_ids <- unique(unlist(sapply(
    espresso_list, function(df) {df$transcript_ID}
)))
espresso_tpm <- sapply(espresso_list, function(df) {
    tpm_values <- df$sample
    names(tpm_values) <- df$transcript_ID
    tpm_values <- tpm_values[espresso_tx_ids]
    tpm_values[is.na(tpm_values)] <- 0
    return(tpm_values)
})
rownames(espresso_tpm) <- espresso_tx_ids
colnames(espresso_tpm) <- sample_ids
espresso_tpm <- t(t(espresso_tpm) / colSums(espresso_tpm) * 1e6)
espresso_mean_rel_diff <- calculate_mean_rel_diff(espresso_tpm)
```

# Results

Results table:
```{r, message = FALSE, warning = FALSE}
results_df <- as.data.frame(cbind(
    Isosceles = isosceles_mean_rel_diff,
    IsoQuant = isoquant_mean_rel_diff,
    Bambu = bambu_mean_rel_diff,
    Flair = flair_mean_rel_diff,
    NanoCount = nanocount_mean_rel_diff,
    LIQA = liqa_mean_rel_diff,
    ESPRESSO = espresso_mean_rel_diff
))
results_df
```

Prepare relative difference data:
```{r, message = FALSE, warning = FALSE}
results_df$platform <- rownames(results_df)
results_df <- pivot_longer(
    results_df, c(-platform), 
    names_to = "tool", values_to = "mean_rel_diff"
)
results_df$tool <- fct_relevel(results_df$tool, "Isosceles")
results_df$tool <- fct_rev(results_df$tool)
results_df$platform <- factor(results_df$platform,
                              levels = c("PP", "MM", "MP"))
```

Bulk mean relative difference plot:
```{r, message = FALSE, warning = FALSE}
ggplot(results_df,
       mapping = aes(x = mean_rel_diff,
                     y = tool,
                     fill = tool)) +
    geom_col() +
    facet_grid(platform ~ .) +
    scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                 ESPRESSO = "lightsalmon",
                                 Flair = "mediumpurple",
                                 Isosceles = "black",
                                 IsoQuant = "gold",
                                 LIQA = "deeppink",
                                 NanoCount = "cornflowerblue")) +
        lims(x = c(0, 1)) +
        labs(
           x = "Mean rel. diff.",
           y = ""
        ) +
        theme_bw() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank())
```

# Session Info

```{r}
sessionInfo()
```
