---
title: "Benchmarking Isosceles on Pacbio, Nanopore and Illumina GM12878 data"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
library(Matrix)
library(SingleCellExperiment)
```

Global parameters:
```{r}
result_dir <- file.path("..", "pacbio_analysis", "report_data")
illumina_sample_id <- "ENCFF485OUK"
pacbio_sample_id <- "ENCFF450VAU"
ont_sample_id <- "NA12878_ONT"
min_tpm <- 1
```

Helper functions:
```{r}
mean_rel_diff <- function(x, y) {
    selector <- (x != 0) & (y != 0)
    x <- x[selector]
    y <- y[selector]
    mean(abs(x - y) / ((x + y) / 2))
}
```

Prepare Isosceles transcript ID map data:
```{r, message = FALSE, warning = FALSE}
se <- readRDS(file.path(
        result_dir, glue("isosceles_{pacbio_sample_id}_se_transcript.rds")
))
tx_df <- rowData(se)[, c("transcript_id", "compatible_tx")]
tx_df <- as.data.frame(tx_df)
rownames(tx_df) <- NULL
tx_df$compatible_tx <- strsplit(tx_df$compatible_tx, ",")
tx_df <- unchop(tx_df, compatible_tx)
tx2isosceles <- setNames(tx_df$transcript_id, tx_df$compatible_tx)
all_isosceles_ids <- sort(unique(tx2isosceles))
all_tx_ids <- sort(names(tx2isosceles))
```

Prepare Illumina TPM data:
```{r, message = FALSE, warning = FALSE}
illumina_df <- read_delim(file.path(
    result_dir, glue("illumina_{illumina_sample_id}.tsv")
))
illumina_df <- filter(illumina_df, grepl("^ENST", transcript_id), TPM > min_tpm)
illumina_tpm <- illumina_df$TPM
names(illumina_tpm) <- sapply(strsplit(illumina_df$transcript_id, "\\."), "[", 1)
illumina_tpm <- illumina_tpm[names(illumina_tpm) %in% all_tx_ids]
illumina_tpm <- illumina_tpm / sum(illumina_tpm) * 1e6
illumina_tx_ids <- names(illumina_tpm)
illumina_isosceles_tpm <- tapply(
    illumina_tpm, tx2isosceles[names(illumina_tpm)], sum
)
illumina_isosceles_tpm <- illumina_isosceles_tpm /
    sum(illumina_isosceles_tpm) * 1e6
illumina_isosceles_tx_ids <- names(illumina_isosceles_tpm)
```

Prepare Isosceles TPM data:
```{r, message = FALSE, warning = FALSE}
se_pacbio <- readRDS(file.path(
    result_dir, glue("isosceles_{pacbio_sample_id}_se_transcript.rds")
))
isosceles_pacbio_tpm <- assay(se_pacbio, "tpm")[, 1]
isosceles_pacbio_tpm <- isosceles_pacbio_tpm[illumina_isosceles_tx_ids]
isosceles_pacbio_tpm <- isosceles_pacbio_tpm / sum(isosceles_pacbio_tpm) * 1e6
se_ont <- readRDS(file.path(
    result_dir, glue("isosceles_{ont_sample_id}_se_transcript.rds")
))
isosceles_ont_tpm <- assay(se_ont, "tpm")[, 1]
isosceles_ont_tpm <- isosceles_ont_tpm[illumina_isosceles_tx_ids]
isosceles_ont_tpm <- isosceles_ont_tpm / sum(isosceles_ont_tpm) * 1e6
```

Prepare IsoQuant TPM data:
```{r, message = FALSE, warning = FALSE}
isoquant_pacbio_df <- read_delim(file.path(
    result_dir, glue("isoquant_{pacbio_sample_id}_transcript_tpm.tsv")
))
isoquant_pacbio_tpm <- isoquant_pacbio_df$TPM
names(isoquant_pacbio_tpm) <- isoquant_pacbio_df[, 1, drop = TRUE]
isoquant_pacbio_tpm <- isoquant_pacbio_tpm[illumina_tx_ids]
isoquant_pacbio_tpm[is.na(isoquant_pacbio_tpm)] <- 0
names(isoquant_pacbio_tpm) <- illumina_tx_ids
isoquant_pacbio_tpm <- isoquant_pacbio_tpm / sum(isoquant_pacbio_tpm) * 1e6
isoquant_ont_df <- read_delim(file.path(
    result_dir, glue("isoquant_{ont_sample_id}_transcript_tpm.tsv")
))
isoquant_ont_tpm <- isoquant_ont_df$TPM
names(isoquant_ont_tpm) <- isoquant_ont_df[, 1, drop = TRUE]
isoquant_ont_tpm <- isoquant_ont_tpm[illumina_tx_ids]
isoquant_ont_tpm[is.na(isoquant_ont_tpm)] <- 0
names(isoquant_ont_tpm) <- illumina_tx_ids
isoquant_ont_tpm <- isoquant_ont_tpm / sum(isoquant_ont_tpm) * 1e6
```

Prepare bambu TPM data:
```{r, message = FALSE, warning = FALSE}
bambu_pacbio_df <- read_delim(file.path(
    result_dir, glue("bambu_{pacbio_sample_id}_counts_transcript.txt")
))
bambu_pacbio_tpm <- bambu_pacbio_df[, 3, drop = TRUE]
names(bambu_pacbio_tpm) <- bambu_pacbio_df$TXNAME
bambu_pacbio_tpm <- bambu_pacbio_tpm / sum(bambu_pacbio_tpm) * 1e6
bambu_pacbio_tpm <- bambu_pacbio_tpm[illumina_tx_ids]
bambu_pacbio_tpm <- bambu_pacbio_tpm / sum(bambu_pacbio_tpm) * 1e6
bambu_ont_df <- read_delim(file.path(
    result_dir, glue("bambu_{ont_sample_id}_counts_transcript.txt")
))
bambu_ont_tpm <- bambu_ont_df[, 3, drop = TRUE]
names(bambu_ont_tpm) <- bambu_ont_df$TXNAME
bambu_ont_tpm <- bambu_ont_tpm / sum(bambu_ont_tpm) * 1e6
bambu_ont_tpm <- bambu_ont_tpm[illumina_tx_ids]
bambu_ont_tpm <- bambu_ont_tpm / sum(bambu_ont_tpm) * 1e6
```

# Results

Prepare Spearman correlation data:
```{r, message = FALSE, warning = FALSE}
spearman_cor_df <- data.frame(
    tool = c("Isosceles", "IsoQuant", "Bambu"),
    pacbio_illumina = c(
        cor(isosceles_pacbio_tpm, illumina_isosceles_tpm, method = "spearman"),
        cor(isoquant_pacbio_tpm, illumina_tpm, method = "spearman"),
        cor(bambu_pacbio_tpm, illumina_tpm, method = "spearman")
    ),
    ont_illumina = c(
        cor(isosceles_ont_tpm, illumina_isosceles_tpm, method = "spearman"),
        cor(isoquant_ont_tpm, illumina_tpm, method = "spearman"),
        cor(bambu_ont_tpm, illumina_tpm, method = "spearman")
    ),
    pacbio_ont = c(
        cor(isosceles_pacbio_tpm, isosceles_ont_tpm, method = "spearman"),
        cor(isoquant_pacbio_tpm, isoquant_ont_tpm, method = "spearman"),
        cor(bambu_pacbio_tpm, bambu_ont_tpm, method = "spearman")
    )
)
```

Spearman correlation data:
```{r, message = FALSE, warning = FALSE}
spearman_cor_df
```

Plot Spearman correlation (Pacbio vs Illumina):
```{r, message = FALSE, warning = FALSE}
spearman_cor_df$value <- spearman_cor_df$pacbio_illumina
spearman_cor_df$tool <- fct_reorder(spearman_cor_df$tool,
                                    spearman_cor_df$value)
ggplot(spearman_cor_df, mapping = aes(x = value,
                                      y = tool,
                                      fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1)) +
        labs(
            x = "Correlation value",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

Plot Spearman correlation (Nanopore vs Illumina):
```{r, message = FALSE, warning = FALSE}
spearman_cor_df$value <- spearman_cor_df$ont_illumina
spearman_cor_df$tool <- fct_reorder(spearman_cor_df$tool,
                                    spearman_cor_df$value)
ggplot(spearman_cor_df, mapping = aes(x = value,
                                      y = tool,
                                      fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1)) +
        labs(
            x = "Correlation value",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

Plot Spearman correlation (Pacbio vs Nanopore):
```{r, message = FALSE, warning = FALSE}
spearman_cor_df$value <- spearman_cor_df$pacbio_ont
spearman_cor_df$tool <- fct_reorder(spearman_cor_df$tool,
                                    spearman_cor_df$value)
ggplot(spearman_cor_df, mapping = aes(x = value,
                                      y = tool,
                                      fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1)) +
        labs(
            x = "Correlation value",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

Prepare mean relative difference data:
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df <- data.frame(
    tool = c("Isosceles", "IsoQuant", "Bambu"),
    pacbio_illumina = c(
        mean_rel_diff(isosceles_pacbio_tpm, illumina_isosceles_tpm),
        mean_rel_diff(isoquant_pacbio_tpm, illumina_tpm),
        mean_rel_diff(bambu_pacbio_tpm, illumina_tpm)
    ),
    ont_illumina = c(
        mean_rel_diff(isosceles_ont_tpm, illumina_isosceles_tpm),
        mean_rel_diff(isoquant_ont_tpm, illumina_tpm),
        mean_rel_diff(bambu_ont_tpm, illumina_tpm)
    ),
    pacbio_ont = c(
        mean_rel_diff(isosceles_pacbio_tpm, isosceles_ont_tpm),
        mean_rel_diff(isoquant_pacbio_tpm, isoquant_ont_tpm),
        mean_rel_diff(bambu_pacbio_tpm, bambu_ont_tpm)
    )
)
```

Mean relative difference data:
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df
```

Plot mean relative difference (Pacbio vs Illumina):
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df$value <- mean_rel_diff_df$pacbio_illumina
mean_rel_diff_df$tool <- fct_reorder(mean_rel_diff_df$tool,
                                     mean_rel_diff_df$value)
ggplot(mean_rel_diff_df, mapping = aes(x = value,
                                       y = tool,
                                       fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1.2)) +
        labs(
            x = "Mean rel. diff.",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

Plot mean relative difference (Nanopore vs Illumina):
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df$value <- mean_rel_diff_df$ont_illumina
mean_rel_diff_df$tool <- fct_reorder(mean_rel_diff_df$tool,
                                     mean_rel_diff_df$value)
ggplot(mean_rel_diff_df, mapping = aes(x = value,
                                       y = tool,
                                       fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1.2)) +
        labs(
            x = "Mean rel. diff.",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

Plot mean relative difference (Pacbio vs Nanopore):
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df$value <- mean_rel_diff_df$pacbio_ont
mean_rel_diff_df$tool <- fct_reorder(mean_rel_diff_df$tool,
                                     mean_rel_diff_df$value)
ggplot(mean_rel_diff_df, mapping = aes(x = value,
                                       y = tool,
                                       fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            hjust = -0.4
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     Isosceles = "black",
                                     IsoQuant = "gold")) +
        lims(x = c(0, 1.2)) +
        labs(
            x = "Mean rel. diff.",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
```

# Session Info

```{r}
sessionInfo()
```
