---
title: "Benchmarking Isosceles on Sequin data"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
library(SingleCellExperiment)
library(patchwork)
```

Global parameters:
```{r}
result_dir <- file.path("..", "sequin_analysis", "report_data")
sample_ids <- c("mixA_ont", "mixB_ont")
```

Helper functions:
```{r}
mean_rel_diff <- function(x, y) {
    mean(abs(x - y) / ((x + y) / 2))
}
```

Prepare reference TPM data:
```{r, message = FALSE, warning = FALSE}
ref_tx_df <- read_delim(file.path(
    result_dir, "sequin_transcripts.tsv"
))
ref_tx_df$MIX_A <- ref_tx_df$MIX_A / sum(ref_tx_df$MIX_A) * 1e6
ref_tx_df$MIX_B <- ref_tx_df$MIX_B / sum(ref_tx_df$MIX_B) * 1e6
```

Prepare Isosceles TPM data:
```{r, message = FALSE, warning = FALSE}
isosceles_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    se_transcript <- readRDS(file.path(
        result_dir, glue("isosceles_{sample_id}_se_transcript.rds")
    ))
    tx_df <- as.data.frame(rowData(se_transcript))
    tpm_values <- assay(se_transcript, "tpm")[, 1]
    names(tpm_values) <- tx_df$compatible_tx
    tpm_values <- tpm_values[ref_tx_df$NAME]
    return(tpm_values)
}))
```

Prepare IsoQuant TPM data:
```{r, message = FALSE, warning = FALSE}
isoquant_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("isoquant_{sample_id}_transcript_tpm.tsv")
    ))
    tpm_values <- tx_df$TPM
    names(tpm_values) <- tx_df$`#feature_id`
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values[is.na(tpm_values)] <- 0
    names(tpm_values) <- ref_tx_df$NAME
    return(tpm_values)
}))
```

Prepare bambu TPM data:
```{r, message = FALSE, warning = FALSE}
bambu_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("bambu_{sample_id}_counts_transcript.txt")
    ))
    tpm_values <- tx_df[, 3, drop = TRUE]
    names(tpm_values) <- tx_df$TXNAME
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values <- tpm_values / sum(tpm_values) * 1e6
    return(tpm_values)
}))
```

Prepare Flair TPM data:
```{r, message = FALSE, warning = FALSE}
flair_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("flair_{sample_id}_counts_matrix.tsv")
    ))
    tx_df <- tx_df[grepl("^R\\d+", tx_df$ids),]
    tx_df$ids <- sapply(
        strsplit(tx_df$ids, "_"),
        function(x) {paste0(x[1:3], collapse = "_")}
    )
    tx_df <- tx_df[!grepl("-", tx_df$ids),]
    tpm_values <- tx_df$sample_condition_batch
    names(tpm_values) <- tx_df$ids
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values[is.na(tpm_values)] <- 0
    names(tpm_values) <- ref_tx_df$NAME
    tpm_values <- tpm_values / sum(tpm_values) * 1e6
    return(tpm_values)
}))
```

Prepare NanoCount TPM data:
```{r, message = FALSE, warning = FALSE}
nanocount_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("nanocount_{sample_id}.tsv")
    ))
    tpm_values <- tx_df$tpm
    names(tpm_values) <- tx_df$transcript_name
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values[is.na(tpm_values)] <- 0
    names(tpm_values) <- ref_tx_df$NAME
    return(tpm_values)
}))
```

Prepare LIQA TPM data:
```{r, message = FALSE, warning = FALSE}
liqa_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("liqa_{sample_id}.tsv")
    ))
    tpm_values <- tx_df$ReadPerGene_corrected
    names(tpm_values) <- tx_df$IsoformName
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values[is.na(tpm_values)] <- 0
    names(tpm_values) <- ref_tx_df$NAME
    tpm_values <- tpm_values / sum(tpm_values) * 1e6
    return(tpm_values)
}))
```

Prepare ESPRESSO TPM data:
```{r, message = FALSE, warning = FALSE}
espresso_tpm_df <- as.data.frame(sapply(sample_ids, function(sample_id) {
    tx_df <- read_delim(file.path(
        result_dir, glue("espresso_{sample_id}_abundance.esp")
    ))
    tpm_values <- tx_df$sample
    names(tpm_values) <- tx_df$transcript_ID
    tpm_values <- tpm_values[ref_tx_df$NAME]
    tpm_values[is.na(tpm_values)] <- 0
    names(tpm_values) <- ref_tx_df$NAME
    tpm_values <- tpm_values / sum(tpm_values) * 1e6
    return(tpm_values)
}))
```

# Results

Prepare Spearman correlation data:
```{r, message = FALSE, warning = FALSE}
tools <- c("Isosceles", "IsoQuant", "Bambu", "Flair", "NanoCount", "LIQA",
           "ESPRESSO")
spearman_cor_df <- data.frame(
    tool = rep(tools, each = 3),
    mix = rep(c("mix_A", "mix_B", "both"), 7),
    value = c(
        cor(isosceles_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(isosceles_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(isosceles_tpm_df$mixA_ont, isosceles_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(isoquant_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(isoquant_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(isoquant_tpm_df$mixA_ont, isoquant_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(bambu_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(bambu_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(bambu_tpm_df$mixA_ont, bambu_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(flair_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(flair_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(flair_tpm_df$mixA_ont, flair_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(nanocount_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(nanocount_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(nanocount_tpm_df$mixA_ont, nanocount_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(liqa_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(liqa_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(liqa_tpm_df$mixA_ont, liqa_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman"),
        cor(espresso_tpm_df$mixA_ont, ref_tx_df$MIX_A, method = "spearman"),
        cor(espresso_tpm_df$mixB_ont, ref_tx_df$MIX_B, method = "spearman"),
        cor(c(espresso_tpm_df$mixA_ont, espresso_tpm_df$mixB_ont),
            c(ref_tx_df$MIX_A, ref_tx_df$MIX_B), method = "spearman")
    )
)
```

Spearman correlation data:
```{r, message = FALSE, warning = FALSE}
spearman_cor_df
```

Prepare mean relative difference data:
```{r, message = FALSE, warning = FALSE}
tools <- c("Isosceles", "IsoQuant", "Bambu", "Flair", "NanoCount", "LIQA",
           "ESPRESSO")
mean_rel_diff_df <- data.frame(
    tool = rep(tools, each = 3),
    mix = rep(c("mix_A", "mix_B", "both"), 7),
    value = c(
        mean_rel_diff(isosceles_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(isosceles_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(isosceles_tpm_df$mixA_ont, isosceles_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(isoquant_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(isoquant_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(isoquant_tpm_df$mixA_ont, isoquant_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(bambu_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(bambu_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(bambu_tpm_df$mixA_ont, bambu_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(flair_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(flair_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(flair_tpm_df$mixA_ont, flair_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(nanocount_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(nanocount_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(nanocount_tpm_df$mixA_ont, nanocount_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(liqa_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(liqa_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(liqa_tpm_df$mixA_ont, liqa_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B)),
        mean_rel_diff(espresso_tpm_df$mixA_ont, ref_tx_df$MIX_A),
        mean_rel_diff(espresso_tpm_df$mixB_ont, ref_tx_df$MIX_B),
        mean_rel_diff(c(espresso_tpm_df$mixA_ont, espresso_tpm_df$mixB_ont),
                      c(ref_tx_df$MIX_A, ref_tx_df$MIX_B))
    )
)
```

Mean relative difference data:
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_df
```

Helper function for plotting Spearman correlation data:
```{r, message = FALSE, warning = FALSE}
plot_spearman_cor <- function(mix_value, title = FALSE, max_x = 1) {
    plot_df <- spearman_cor_df %>%
        filter(mix == mix_value)
    plot_df$tool <- fct_reorder(plot_df$tool, plot_df$value)
    bar_plot <- ggplot(plot_df, mapping = aes(x = value,
                                              y = tool,
                                              fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            nudge_x = 0.03,
            hjust = "outward"
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     ESPRESSO = "lightsalmon",
                                     Flair = "mediumpurple",
                                     Isosceles = "black",
                                     IsoQuant = "gold",
                                     LIQA = "deeppink",
                                     NanoCount = "cornflowerblue")) +
        lims(x = c(0, max_x)) +
        labs(
            x = "Correlation value",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
    if (title) {
        bar_plot <- bar_plot +
            labs(title = mix_value)
    }
    return(bar_plot)
}
```

Plot Spearman correlation (mix_A):
```{r, message = FALSE, warning = FALSE}
plot_spearman_cor("mix_A", max_x = 1.1)
```

Plot Spearman correlation (mix_B):
```{r, message = FALSE, warning = FALSE}
plot_spearman_cor("mix_B", max_x = 1.1)
```

Plot Spearman correlation (both mixes):
```{r, message = FALSE, warning = FALSE}
plot_spearman_cor("both", max_x = 1.1)
```

Plot Spearman correlation (mixA & mix_B):
```{r, message = FALSE, warning = FALSE}
plot_spearman_cor("mix_A", title = TRUE, max_x = 1.25) +
    plot_spearman_cor("mix_B", title = TRUE, max_x = 1.25)
```

Helper function for plotting mean relative difference data:
```{r, message = FALSE, warning = FALSE}
plot_mean_rel_diff <- function(mix_value, title = FALSE, max_x = 2) {
    plot_df <- mean_rel_diff_df %>%
        filter(mix == mix_value)
    plot_df$tool <- fct_reorder(plot_df$tool, plot_df$value)
    plot_df$tool <- fct_rev(plot_df$tool)
    bar_plot <- ggplot(plot_df, mapping = aes(x = value,
                                              y = tool,
                                              fill = tool)) +
        geom_col() +
        geom_text(
            aes(label = format(round(value, 2), nsmall = 2)),
            size = 3,
            nudge_x = 0.03,
            hjust = "outward"
        ) +
        scale_fill_manual(values = c(Bambu = "mediumseagreen",
                                     ESPRESSO = "lightsalmon",
                                     Flair = "mediumpurple",
                                     Isosceles = "black",
                                     IsoQuant = "gold",
                                     LIQA = "deeppink",
                                     NanoCount = "cornflowerblue")) +
        lims(x = c(0, max_x)) +
        labs(
            x = "Mean rel. diff.",
            y = ""
        ) +
        theme_classic() +
        theme(legend.position = "none",
              axis.text = element_text(size = 7),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 9),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank())
    if (title) {
        bar_plot <- bar_plot +
            labs(title = mix_value)
    }
    return(bar_plot)
}
```

Plot mean relative difference (mix_A):
```{r, message = FALSE, warning = FALSE}
plot_mean_rel_diff("mix_A", max_x = 1.3)
```

Plot mean relative difference (mix_B):
```{r, message = FALSE, warning = FALSE}
plot_mean_rel_diff("mix_B", max_x = 1.3)
```

Plot mean relative difference (both mixes):
```{r, message = FALSE, warning = FALSE}
plot_mean_rel_diff("both", max_x = 1.3)
```

Plot mean relative difference (mixA & mix_B):
```{r, message = FALSE, warning = FALSE}
plot_mean_rel_diff("mix_A", title = TRUE, max_x = 1.47) +
    plot_mean_rel_diff("mix_B", title = TRUE, max_x = 1.47)
```

# Session Info

```{r}
sessionInfo()
```
