---
title: "Simulated ovarian cell line bulk RNA-Seq and scRNA-Seq data benchmarks (top highly variable transcripts results summary)"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
```

Global parameters:
```{r, message = FALSE, warning = FALSE}
csv_dir <- "simulated_ovarian_bulk_sc_benchmarks_hvts_data"
top_n_hvts_values <- c(500, 1000, 2000, 4000, 6000, 10000)
```

Prepare plot data:
```{r, message = FALSE, warning = FALSE}
plot_df_list <- lapply(top_n_hvts_values, function(top_n_hvts) {
    plot_df <- read.csv(file.path(
        csv_dir, glue("data_{top_n_hvts}_hvts.csv")
    ), row.names = 1)
    plot_df <- as.data.frame(t(plot_df))
    plot_df$tool <- rownames(plot_df)
    rownames(plot_df) <- NULL
    plot_df$top_n_hvts <- top_n_hvts
    plot_df <- plot_df %>%
        select("tool", "top_n_hvts", "corr_spearman",
               "corr_spearman_upper_ci_95", "corr_spearman_lower_ci_95",
                "corr_spearman_upper_ci_90", "corr_spearman_lower_ci_90",
               "corr_spearman_upper_ci_68", "corr_spearman_lower_ci_68",
               "mean_rel_diff",
               "mean_rel_diff_upper_ci_95", "mean_rel_diff_lower_ci_95",
               "mean_rel_diff_upper_ci_90", "mean_rel_diff_lower_ci_90",
               "mean_rel_diff_upper_ci_68", "mean_rel_diff_lower_ci_68")
    return(plot_df)
})
plot_df <- do.call(rbind, plot_df_list)
cor_spearman_hvts_fc_df <- plot_df %>%
    select("tool", "top_n_hvts", "corr_spearman") %>%
    pivot_wider(names_from = tool, values_from = corr_spearman) %>%
    mutate(
        IsoQuant = Isosceles / IsoQuant,
        FLAMES = Isosceles / FLAMES,
        Sicelore = Isosceles / Sicelore
    ) %>%
    select(-"Isosceles")
cor_spearman_fc_df <- plot_df %>%
    select("tool", "top_n_hvts", "corr_spearman") %>%
    pivot_wider(names_from = tool, values_from = corr_spearman) %>%
    transmute(
        IsoQuant = Isosceles / IsoQuant,
        FLAMES = Isosceles / FLAMES,
        Sicelore = Isosceles / Sicelore
    )
cor_spearman_fc_df <- data.frame(
    median_fc = apply(cor_spearman_fc_df, 2, median),
    max_fc = apply(cor_spearman_fc_df, 2, max)
)
mean_rel_diff_hvts_fc_df <- plot_df %>%
    select("tool", "top_n_hvts", "mean_rel_diff") %>%
    pivot_wider(names_from = tool, values_from = mean_rel_diff) %>%
    mutate(
        IsoQuant = Isosceles / IsoQuant,
        FLAMES = Isosceles / FLAMES,
        Sicelore = Isosceles / Sicelore
    ) %>%
    select(-"Isosceles")
mean_rel_diff_fc_df <- plot_df %>%
    select("tool", "top_n_hvts", "mean_rel_diff") %>%
    pivot_wider(names_from = tool, values_from = mean_rel_diff) %>%
    transmute(
        IsoQuant = Isosceles / IsoQuant,
        FLAMES = Isosceles / FLAMES,
        Sicelore = Isosceles / Sicelore
    )
mean_rel_diff_fc_df <- data.frame(
    median_fc = apply(mean_rel_diff_fc_df, 2, median),
    max_fc = apply(mean_rel_diff_fc_df, 2, max)
)
test_df_list <- lapply(top_n_hvts_values, function(top_n_hvts) {
    test_df <- read.csv(file.path(
        csv_dir, glue("test_values_{top_n_hvts}_hvts.csv")
    ))
    test_df$top_n_hvts <- top_n_hvts
    return(test_df)
})
test_df <- do.call(rbind, test_df_list)
corr_spearman_isosceles <- filter(test_df, tool == "Isosceles") %>%
    pull(corr_spearman)
mean_rel_diff_isosceles <- filter(test_df, tool == "Isosceles") %>%
    pull(mean_rel_diff)
wilcox_pvalue <- function(x, y) {
    wilcox.test(x, y, paired = TRUE)$p.value
}
test_df <- test_df %>%
    filter(tool != "Isosceles") %>%
    group_by(tool) %>%
    summarise(
        corr_spearman_pvalue = wilcox_pvalue(corr_spearman,
                                             corr_spearman_isosceles),
        mean_rel_diff_pvalue = wilcox_pvalue(mean_rel_diff,
                                             mean_rel_diff_isosceles)
    ) %>%
    as.data.frame()
rownames(test_df) <- test_df$tool
test_df$tool <- NULL
test_df <- test_df[c("IsoQuant", "FLAMES", "Sicelore"),]
cor_spearman_fc_df$pvalue <- test_df$corr_spearman_pvalue
mean_rel_diff_fc_df$pvalue <- test_df$mean_rel_diff_pvalue
```

# Results

Matched - decoy fold-change values (Spearman correlation):
```{r, message = FALSE, warning = FALSE}
cor_spearman_hvts_fc_df
```

Matched - decoy fold-change statistics (Spearman correlation):
```{r, message = FALSE, warning = FALSE}
cor_spearman_fc_df
```

Matched - decoy fold-change values (mean relative difference):
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_hvts_fc_df
```

Matched - decoy fold-change statistics (mean relative difference):
```{r, message = FALSE, warning = FALSE}
mean_rel_diff_fc_df
```

Matched - decoy plot (Spearman correlation, standard error interval):
```{r, message = FALSE, warning = FALSE}
ggplot(plot_df, mapping = aes(x = top_n_hvts)) +
    geom_line(mapping = aes(y = corr_spearman,
                            col = tool),
              size = 1) +
    geom_ribbon(mapping = aes(ymin = corr_spearman_lower_ci_68,
                              ymax = corr_spearman_upper_ci_68,
                              fill = tool),
                alpha = 0.1) +
    scale_color_manual(values = c(Isosceles = "black",
                                  IsoQuant = "gold",
                                  FLAMES = "tomato",
                                  Sicelore = "royalblue")) +
    scale_fill_manual(values = c(Isosceles = "black",
                                 IsoQuant = "gold",
                                 FLAMES = "tomato",
                                 Sicelore = "royalblue")) +
    scale_x_continuous(
        breaks = c(500, 1000, 2000, 4000, 6000, 10000),
        labels = c("500", "1k", "2k", "4k", "6k", "10k")
    ) +
    labs(
        x = "Top N variable transcripts",
        y = "Matched - decoy (correlation)"
    ) +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.title = element_blank(),
          axis.text.x = element_text(size = 7, hjust = 0.85))
```

Matched - decoy plot (mean relative difference, standard error interval):
```{r, message = FALSE, warning = FALSE}
ggplot(plot_df, mapping = aes(x = top_n_hvts)) +
    geom_line(mapping = aes(y = mean_rel_diff,
                            col = tool),
              size = 1) +
    geom_ribbon(mapping = aes(ymin = mean_rel_diff_lower_ci_68,
                              ymax = mean_rel_diff_upper_ci_68,
                              fill = tool),
                alpha = 0.1) +
    scale_color_manual(values = c(Isosceles = "black",
                                  IsoQuant = "gold",
                                  FLAMES = "tomato",
                                  Sicelore = "royalblue")) +
    scale_fill_manual(values = c(Isosceles = "black",
                                 IsoQuant = "gold",
                                 FLAMES = "tomato",
                                 Sicelore = "royalblue")) +
    scale_x_continuous(
        breaks = c(500, 1000, 2000, 4000, 6000, 10000),
        labels = c("500", "1k", "2k", "4k", "6k", "10k")
    ) +
    labs(
        x = "Top N variable transcripts",
        y = "Matched - decoy (mean rel. diff.)"
    ) +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.title = element_blank(),
          axis.text.x = element_text(size = 7, hjust = 0.85))
```

# Session Info

```{r}
sessionInfo()
```
