---
title: "Simulated scRNA-Seq data benchmarks"
output:
  html_document:
    highlight: pygments
    df_print: paged
    toc: true
    toc_depth: 2
---

# Preparing the data

Load the required packages:
```{r, message = FALSE}
library(tidyverse)
library(glue)
library(Matrix)
library(Hmisc)
library(SingleCellExperiment)
```

Global parameters:
```{r}
result_dir <- file.path("..", "simulated_sc_analysis", "report_data")
```

Helper functions:
```{r}
fill_missing_matrix <- function(x, all_rownames) {
    missing_rownames <- setdiff(all_rownames, rownames(x))
    missing_matrix <- as.matrix(Matrix(
        0, nrow = length(missing_rownames), ncol = ncol(x)
    ))
    rownames(missing_matrix) <- missing_rownames
    full_matrix <- rbind(x, missing_matrix)
    full_matrix <- full_matrix[all_rownames,]
    return(full_matrix)
}
median_rel_diff <- function(x, y) {
    median(abs(x - y) / ((x + y) / 2))
}
mean_rel_diff <- function(x, y) {
    mean(abs(x - y) / ((x + y) / 2))
}
```

Prepare transcript class (i.e. unique intron structure) data:
```{r, message = FALSE, warning = FALSE}
se_transcript <- readRDS(file.path(
    result_dir, "isosceles_se_transcript.rds"
))
tx_df <- rowData(se_transcript)[, c("intron_positions", "compatible_tx")]
tx_df <- as.data.frame(tx_df)
rownames(tx_df) <- NULL
tx_df$compatible_tx <- strsplit(tx_df$compatible_tx, ",")
tx_df <- unchop(tx_df, compatible_tx)
tx_df$class_id <- as.numeric(as.factor(tx_df$intron_positions))
tx2class <- setNames(tx_df$class_id, tx_df$compatible_tx)
all_class_ids <- as.character(sort(unique(tx2class)))
```

Prepare reference TPM data:
```{r, message = FALSE, warning = FALSE}
ref_tpm_df <- read.delim(file.path("..", "reference_data",
                                   "benchmark_transcript_expression.tab"))
ref_class_ids <- unname(tx2class[ref_tpm_df$target_id])
ref_tpm <- unname(tapply(ref_tpm_df$tpm, ref_class_ids, sum))
```

Prepare reference TPM quantile data:
```{r, message = FALSE, warning = FALSE}
ref_quantile <- ceiling(seq_along(ref_tpm) * 33/length(ref_tpm)) * 99/33
ref_quantile <- ref_quantile[rank(ref_tpm, ties.method = 'first')]
```

Prepare Isosceles TPM data:
```{r, message = FALSE, warning = FALSE}
isosceles_tpm <- assay(se_transcript, "tpm")
isosceles_class_ids <- unname(tx2class[
    sapply(strsplit(rowData(se_transcript)$compatible_tx, ","), "[", 1)
])
isosceles_tpm <- rowsum(assay(se_transcript, "tpm"), isosceles_class_ids)
```

Prepare Isosceles pseudobulk TPM data:
```{r, message = FALSE, warning = FALSE}
se_pseudobulk_transcript <- readRDS(file.path(
    result_dir, "isosceles_se_pseudobulk_transcript.rds"
))
isosceles_pseudobulk_tpm <- assay(se_pseudobulk_transcript, "tpm")[, 1]
isosceles_pseudobulk_class_ids <- unname(tx2class[
    sapply(strsplit(rowData(se_pseudobulk_transcript)$compatible_tx,","),
           "[", 1)
])
isosceles_pseudobulk_tpm <- as.numeric(unname(tapply(
    isosceles_pseudobulk_tpm, isosceles_pseudobulk_class_ids, sum
)))
```

Prepare IsoQuant TPM data:
```{r, message = FALSE, warning = FALSE}
isoquant_df <- read_delim(file.path(
    result_dir, "isoquant_transcript_grouped_tpm.tsv"
))
isoquant_rowdata <- as.data.frame(isoquant_df[, 1])
isoquant_tpm <- as.matrix(isoquant_df[, c(-1)])
isoquant_tpm <- as(isoquant_tpm, "dgCMatrix")
rownames(isoquant_tpm) <- isoquant_rowdata[[1]]
isoquant_class_ids <- unname(tx2class[rownames(isoquant_tpm)])
isoquant_tpm <- rowsum(isoquant_tpm, isoquant_class_ids)
isoquant_tpm <- fill_missing_matrix(isoquant_tpm, all_class_ids)
```

Prepare IsoQuant pseudobulk TPM data:
```{r, message = FALSE, warning = FALSE}
isoquant_pseudobulk_tpm <- rowSums(isoquant_tpm)
isoquant_pseudobulk_tpm <- isoquant_pseudobulk_tpm /
    sum(isoquant_pseudobulk_tpm) * 1e6
```

Prepare FLAMES TPM data:
```{r, message = FALSE, warning = FALSE}
flames_df <- read_csv(file.path(
    result_dir, "flames_transcript_count.csv.gz"
))
flames_rowdata <- as.data.frame(flames_df[, 1:2])
flames_counts <- as.matrix(flames_df[, c(-1, -2)])
flames_counts <- as(flames_counts, "dgCMatrix")
rownames(flames_counts) <- flames_rowdata$transcript_id
flames_tx_selector <- !grepl("_", rownames(flames_counts))
flames_counts <- flames_counts[flames_tx_selector,]
flames_tpm <- t(t(flames_counts) / colSums(flames_counts) * 1e6)
flames_class_ids <- unname(tx2class[rownames(flames_tpm)])
flames_tpm <- rowsum(flames_tpm, flames_class_ids)
flames_tpm <- fill_missing_matrix(flames_tpm, all_class_ids)
```

Prepare FLAMES pseudobulk TPM data:
```{r, message = FALSE, warning = FALSE}
flames_pseudobulk_tpm <- rowSums(flames_tpm)
flames_pseudobulk_tpm <- flames_pseudobulk_tpm /
    sum(flames_pseudobulk_tpm) * 1e6
```

Prepare Sicelore TPM data:
```{r, message = FALSE, warning = FALSE}
sicelore_df <- read_delim(file.path(
    result_dir, "sicelore_isomatrix.txt"
))
sicelore_rowdata <- as.data.frame(sicelore_df[, 1:3])
sicelore_counts <- as.matrix(sicelore_df[, c(-1, -2, -3)])
sicelore_counts <- as(sicelore_counts, "dgCMatrix")
rownames(sicelore_counts) <- sicelore_rowdata$transcriptId
sicelore_counts <- sicelore_counts[rownames(sicelore_counts) != "undef",]
sicelore_tpm <- t(t(sicelore_counts) / colSums(sicelore_counts) * 1e6)
sicelore_class_ids <- unname(tx2class[rownames(sicelore_tpm)])
sicelore_tpm <- rowsum(sicelore_tpm, sicelore_class_ids)
sicelore_tpm <- fill_missing_matrix(sicelore_tpm, all_class_ids)
```

Prepare Sicelore pseudobulk TPM data:
```{r, message = FALSE, warning = FALSE}
sicelore_pseudobulk_tpm <- rowSums(sicelore_tpm)
sicelore_pseudobulk_tpm <- sicelore_pseudobulk_tpm /
    sum(sicelore_pseudobulk_tpm) * 1e6
```

# Relative difference plots

## Individual cells

```{r, message = FALSE, warning = FALSE, eval=TRUE}
sc_df <- data.frame(
    ref_tpm = ref_tpm,
    ref_quantile = ref_quantile,
    Isosceles = as.numeric(isosceles_tpm),
    IsoQuant = as.numeric(isoquant_tpm),
    FLAMES = as.numeric(flames_tpm),
    Sicelore = as.numeric(sicelore_tpm)
) %>%
    gather(key = "tool", value = "estimated_tpm",
           Isosceles, IsoQuant, FLAMES, Sicelore)
sc_rel_diff_df <- sc_df %>%
    group_by(ref_quantile, tool) %>%
    summarise(
        mean_rel_diff = mean_rel_diff(ref_tpm, estimated_tpm),
        median_rel_diff = median_rel_diff(ref_tpm, estimated_tpm)
    ) %>%
    ungroup()
ggplot(sc_rel_diff_df, mapping = aes(x = ref_quantile,
                                     y = mean_rel_diff,
                                     col = tool)) +
    geom_line(size = 1) +
    scale_color_manual(values = c(Isosceles = "black",
                                  IsoQuant = "gold",
                                  FLAMES = "tomato",
                                  Sicelore = "royalblue")) +
    labs(
        x = "Reference TPM quantile",
        y = "Mean rel. diff.",
        col = ""
    ) +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.title = element_blank())
```

## Pseudobulk data

```{r, message = FALSE, warning = FALSE}
pseudobulk_df <- data.frame(
    ref_tpm = ref_tpm,
    ref_quantile = ref_quantile,
    Isosceles = isosceles_pseudobulk_tpm,
    IsoQuant = isoquant_pseudobulk_tpm,
    FLAMES = flames_pseudobulk_tpm,
    Sicelore = sicelore_pseudobulk_tpm
) %>%
    gather(key = "tool", value = "estimated_tpm",
           Isosceles, IsoQuant, FLAMES, Sicelore)
pseudobulk_rel_diff_df <- pseudobulk_df %>%
    group_by(ref_quantile, tool) %>%
    summarise(
        mean_rel_diff = mean_rel_diff(ref_tpm, estimated_tpm),
        median_rel_diff = median_rel_diff(ref_tpm, estimated_tpm)
    ) %>%
    ungroup()
ggplot(pseudobulk_rel_diff_df,
       mapping = aes(x = ref_quantile,
                     y = mean_rel_diff,
                     col = tool)) +
    geom_line(size = 1) +
    scale_color_manual(values = c(Isosceles = "black",
                                  IsoQuant = "gold",
                                  FLAMES = "tomato",
                                  Sicelore = "royalblue")) +
    labs(
        x = "Reference TPM quantile",
        y = "Mean rel. diff.",
        col = ""
    ) +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.title = element_blank())
```

## Joint plot

```{r, message = FALSE, warning = FALSE}
sc_df <- data.frame(
    ref_tpm = ref_tpm,
    ref_quantile = ref_quantile,
    n_cells = "Single cell",
    Isosceles = as.numeric(isosceles_tpm),
    IsoQuant = as.numeric(isoquant_tpm),
    FLAMES = as.numeric(flames_tpm),
    Sicelore = as.numeric(sicelore_tpm)
)
pseudobulk_df <- data.frame(
    ref_tpm = ref_tpm,
    ref_quantile = ref_quantile,
    n_cells = "Pseudobulk",
    Isosceles = isosceles_pseudobulk_tpm,
    IsoQuant = isoquant_pseudobulk_tpm,
    FLAMES = flames_pseudobulk_tpm,
    Sicelore = sicelore_pseudobulk_tpm
)
plot_data_df <- rbind(sc_df, pseudobulk_df) %>%
    gather(key = "tool", value = "estimated_tpm",
           Isosceles, IsoQuant, FLAMES, Sicelore)
joint_rel_diff_df <- plot_data_df %>%
    group_by(ref_quantile, tool, n_cells) %>%
    summarise(
        mean_rel_diff = mean_rel_diff(ref_tpm, estimated_tpm),
        median_rel_diff = median_rel_diff(ref_tpm, estimated_tpm)
    ) %>%
    ungroup()
ggplot(joint_rel_diff_df,
       mapping = aes(x = ref_quantile,
                     y = mean_rel_diff,
                     col = tool,
                     linetype = n_cells)) +
    geom_line(size = 1) +
    scale_color_manual(values = c(Isosceles = "black",
                                  IsoQuant = "gold",
                                  FLAMES = "tomato",
                                  Sicelore = "royalblue")) +
    scale_linetype_manual(values = c(`Single cell` = "dotted",
                                     `Pseudobulk` = "solid")) +
    labs(
        x = "Reference TPM quantile",
        y = "Mean rel. diff.",
        col = "Tool",
        linetype = "Number of cells"
    ) +
    theme_bw() +
    theme(aspect.ratio = 1,
          legend.title = element_blank())
```

# Session Info

```{r}
sessionInfo()
```
